---
title: "README"
output: html_notebook
---

To install the package, please uncomment and run:
```{r}
install.packages(".", repos = NULL, type = "source")
library(focus.new)
```

# Gaussian


## Pre change parameter known

Here we test the pre-change parameter known recursion, for gaussian change in mean. 
```{r}
theta0 <- 0
set.seed(45)
Y <- c(rnorm(1e3, theta0), rnorm(500, theta0 - 1))

res <- focus_offline(Y, 50, family = "gaussian", theta0 = theta0, args = list(), adp_max_check = F)
plot(res$stat, type = "l")
```
## pre-change unknown
To specify the pre-change unkown we have to set the theta0 parameter to NaN.

```{r}
theta0 <- 0
set.seed(45)
Y <- c(rnorm(1e3, theta0), rnorm(500, theta0 - 1))

res <- focus_offline(Y, 50, family = "gaussian", theta0 = NaN, args = list(), adp_max_check = F)
plot(res$stat, type = "l")
```

# Gamma

## known
Please specify the shape argument in the args list to perform exponential! 

```{r}
theta0 <- 4
shape <- 4
set.seed(42)
Y <- c(rgamma(500, rate = theta0, shape = shape), rgamma(500, rate = theta0 - 1, shape = shape))

# note that I have coded it for change in scale! As in the paper, but this can be changed any time in the interface. 
system.time(res <- focus_offline(Y, 50, family = "gamma", theta0 = 1/theta0, args = list(shape = 4), adp_max_check = F))
plot(res$stat, type = "l")
```

## unknown

```{r}
theta0 <- 4
shape <- 4
set.seed(42)
Y <- c(rgamma(500, rate = theta0, shape = shape), rgamma(500, rate = theta0 - 1, shape = shape))

# note that I have coded it for change in scale! As in the paper, but this can be changed any time in the interface. 
system.time(res <- focus_offline(Y, 50, family = "gamma", theta0 = NaN, args = list(shape = 4), adp_max_check = F))
plot(res$stat, type = "l")
```


# Other distributions

Similarly, we have Bernoulli and Poisson. 

```{r}

### poisson ####
theta0 <- 2
set.seed(42)
Y <- c(rpois(500, theta0), rpois(500, 1.5))

system.time(res <- focus_offline(Y, 50, family = "poisson", theta0 = theta0, args = list(shape = 4), adp_max_check = F))
plot(res$stat, type = "l")



#### bernoulli ####
theta0 <- 0.4
set.seed(45)
Y <- c(rbinom(1e3, 1, theta0), rbinom(1e3, 1, theta0 + .2))
system.time(res <- focus_offline(Y, 50, family = "bernoulli", theta0 = theta0, args = list(), adp_max_check = F))
plot(res$stat, type = "l")
```

# NP-FOCuS
Finally, NP-FOCuS can run as:

```{r}
set.seed(42)
Y <- c(rnorm(5000), rnorm(100, sd = 2))


probs <- computeprobsNPPELT(15)
quants <- qnorm(probs)
theta0 <- pnorm(quants)
system.time(res <- npfocus_offline(Y = Y, threshold = c(90, 15), quantiles = quants))


par(mfrow = c(2, 1))
plot(res$max_stat, type = "l")
plot(res$sum_stat, type = "l")
par(mfrow = c(1, 1))
```



